---
title: "PCAplots"
author: "Lauren E Johnson"
date: "11/29/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


Load the data:
```{r message=FALSE}
load(here::here('data/whole_brain_coords_S1.Rda'))
load(here::here('data/tel_coords_S1.Rda'))
load(here::here('data/dien_coords_S1.Rda'))
load(here::here('data/mes_coords_S1.Rda'))
load(here::here('data/cere_coords_S1.Rda'))
load(here::here('data/tel_coords_S1.Rda'))
load(here::here('data/medob_coords_S1.Rda'))

load(here::here('data/lizard_trait_data.Rda'))

load(here::here('data/lizard_tree.Rda'))


```

Perform a **generalize procrustes analysis (GPA)** with `gpagen` on the coordinate data. GPA aligns all the species coordinates in the most optimal way (e.g., removing scale, orientation, and position) so that we can better compare the shapes of each species to each other.
```{r message=FALSE}
library(geomorph)

wb.gpa <- gpagen(wholebrain, ProcD = T, verbose = T, print.progress = F)

```

Creating a geomorph data frame that contains the Procrustes coordinates and the ecological and reproductive trait data

```{r message=FALSE}

wb.gdf <- geomorph.data.frame(wb.gpa, trait)

```

**Performing a principle components analysis with phylogeny considered (phyloPCA)** on Procrustes shape coordinates and plotting them
```{r}
library(knitr)
library(geomorph)

wb.pca <- gm.prcomp(wb.gpa$coords, phy = lzrd_tree, GLS = TRUE)

#kable(summary(wb.pca))



```


Correlation Matrix
```{r}
library(Hmisc)
library(corrplot)

corrmatrix <- rcorr(as.matrix(trait[,c("maxSVL", "fSVL", "hatchlingSVL", 
                                       "clutch_size", "precip_seasonality", 
                                       "temp_seasonality")]))

# Need to run the GPA first if I want to see if Csize is correlated!!!!!!!!!!!!
mydata.cor <- cor(cbind(wb.gdf$Csize, abs(trait$latitude), (trait[,c("maxSVL", "fSVL","hatchlingSVL", "clutch_size", "precip_seasonality", "temp_seasonality")])), 
                  method = c('pearson'))

mydata.cor

corrmatrix

testRes <- cor.mtest(cbind(wb.gdf$Csize, abs(trait$latitude), (trait[,c("maxSVL", "fSVL", "hatchlingSVL", "clutch_size", "precip_seasonality", "temp_seasonality")])), conf.level = 0.95)

testRes$p

library(RColorBrewer)

corrplot(mydata.cor, method = 'circle', type = 'lower', 
         addCoef.col ='black', number.cex = 0.8, order = 'AOE', diag=FALSE)

#### THIS IS THE GOOD GRAPH!!!! ######
corrplot(mydata.cor, p.mat = testRes$p, method = 'circle', type = 'lower', 
         addCoef.col ='black', number.cex = 0.8, order = 'AOE', 
         diag=FALSE, 
         col=colorRampPalette(c("darkgoldenrod2", "white", "darkcyan"))(100), 
         tl.srt = 45,
         tl.col = 'black')



library(sjPlot)
library(stargazer)


```



Plots
```{r}

# Print the plot to a pdf file
pdf(file = here::here("vignettes/phylopcplot.pdf"), width = 4, height = 4)

plot(wb.pca, phylo = TRUE, main = "Phylo PCA", 
     phylo.par = list(tip.txt.cex = 0.6,
                      node.labels = F,
                      node.cex = 0.5,
                      tip.txt.adj = c(1.1, -0.5)))

dev.off()

library(geomorph)
library(ape)
library(phytools)

plotTree(lzrd_tree, ftype = 'i')
nodelabels(frame = 'circ', bg = 'white', cex = 0.8)



# wb.pca$x #pc score
# 
# plot(wb.pca, time.plot = T, pch = 22, bg = c(rep('red', 5), rep('green', 4)), cex = 2, phylo.par = list(edge.color = 'grey60', edge.width = 1.5, tip.txt.cex = 0.75, node.labels = F, anc.states = F))
# 
# 
# pca3d::pca3d(wb.pca$x)
# 
# 
# # 11/18/2021
# x1 <- wb.pca$x[,1:3]
# str(x1)
# x1
# order(x1[,1])
# order(x1[,2])
# order(x1[,3])

```


brain size plots and traits 

```{r}
plot(wb.gdf$Csize~wb.gdf$precip_seasonality)


df <- data.frame(wb.gdf$Csize, wb.gdf$latitude, wb.gdf$precip_seasonality, wb.gdf$temp_seasonality, wb.gdf$clutch_size, wb.gdf$fSVL, wb.gdf$hatchlingSVL, wb.gdf$substrate, wb.gdf$foraging_mode)

library(ggplot2)
library(hrbrthemes)

# temperature seasonality 
ggplot(df, aes(x=wb.gdf.temp_seasonality, y=wb.gdf.Csize)) +
  geom_point() +
  geom_smooth(method=lm , color="black", fill="lightblue3", se=TRUE) +
  xlab("Temperature Seasonality") + 
  ylab("Centroid Size") + 
  theme_classic()


ggplot(df, aes(x=wb.gdf.temp_seasonality, y=log10(wb.gdf.Csize))) +
  geom_point() +
  geom_smooth(method=lm , color="black", fill="lightblue3", se=TRUE) +
  xlab("Temperature Seasonality") + 
  ylab("log10 Centroid Size") + 
  theme_classic()

# precipitation seasonality 
ggplot(df, aes(x=wb.gdf.precip_seasonality, y=wb.gdf.Csize)) +
  geom_point() +
  geom_smooth(method=lm , color="black", fill="lightblue3", se=TRUE) +
  xlab("Precipitation Seasonality") + 
  ylab("Centroid Size") + 
  theme_classic()

# latitude 
ggplot(df, aes(x=abs(wb.gdf.latitude), y=wb.gdf.Csize)) +
  geom_point() +
  geom_smooth(method=lm , color="black", fill="lightblue3", se=TRUE) +
  xlab("Latitude") + 
  ylab("Centroid Size") + 
  theme_classic()


# clutch size
ggplot(df, aes(x=wb.gdf.clutch_size, y=wb.gdf.Csize)) +
  geom_point() +
  geom_smooth(method=lm , color="black", fill="lightblue3", se=TRUE) +
  xlab("Clutch Size") + 
  ylab("Centroid Size") + 
  theme_classic()


# fsvl

ggplot(df, aes(x=wb.gdf.fSVL, y=wb.gdf.Csize)) +
  geom_point() +
  geom_smooth(method=lm , color="black", fill="lightblue3", se=TRUE) +
  xlab("Female SVL") + 
  ylab("Centroid Size") + 
  theme_classic()

# hatchling svl

ggplot(df, aes(x=wb.gdf.hatchlingSVL, y=wb.gdf.Csize)) +
  geom_point() +
  geom_smooth(method=lm , color="black", fill="lightblue3", se=TRUE) +
  xlab("Hatchling SVL") + 
  ylab("Centroid Size") + 
  theme_classic()
```

boxplots 
```{r}
# substrate

df %>%
  ggplot(aes(x=wb.gdf.substrate, y=wb.gdf.Csize, fill=wb.gdf.substrate)) +
    geom_boxplot() +
    geom_jitter(color="black", size=1, alpha=0.9) +
    theme_classic() +
    theme(
      legend.position="none",
      plot.title = element_text(size=11), 
      axis.text.x = element_text(angle = 60, hjust = 1)
    ) +
    xlab("") + 
    ylab("Centroid Size")

# foraging mode
df %>%
  ggplot(aes(x=wb.gdf.foraging_mode, y=wb.gdf.Csize, fill=wb.gdf.foraging_mode)) +
    geom_boxplot() +
    geom_jitter(color="black", size=1, alpha=0.9) +
    theme_classic() +
    theme(
      legend.position="none",
      plot.title = element_text(size=11), 
      axis.text.x = element_text(angle = 60, hjust = 1)
    ) +
    xlab("") + 
    ylab("Centroid Size")

```

pca 3d plots
```{r}
library(pca3d)
# pca3d(wb.pca$x, group = trait$substrate, legend = "topleft", 
#       show.ellipses = T, ellipse.ci = 0.75)

pca3d(wb.pca$x, group = trait$activity_time, legend = "topleft", 
      show.ellipses = T, ellipse.ci = 0.75)

pca3d(wb.pca$x, group = trait$foraging_mode, legend = "topleft", 
      show.ellipses = T, ellipse.ci = 0.75)

pca3d(wb.pca$x, group = trait$reproductive_mode, legend = "topleft", 
      show.ellipses = T, ellipse.ci = 0.75)

pca2d(wb.pca$x, group = trait$activity_time, legend = "topright", 
      show.ellipses = T, ellipse.ci = 0.75)
box(which = "plot", lty = "solid")

pca2d(wb.pca$x, group = trait$foraging_mode, legend = "topright", 
      show.ellipses = T, ellipse.ci = 0.75)
box(which = "plot", lty = "solid")

pca2d(wb.pca$x, group = trait$reproductive_mode, legend = "topright", 
      show.ellipses = T, ellipse.ci = 0.75)
box(which = "plot", lty = "solid")

```
