---
title: "wireframes"
author: "Lauren E Johnson"
date: "11/22/2021"
output: html_document
---

```{r, setup, echo = FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# setting up hook to save rgl files
library(rgl)
knitr::knit_hooks$set(webgl = hook_webgl)
```

```{r echo=FALSE, include=FALSE, message = FALSE, warning = FALSE, results='hide'}
library(rgl)
setupKnitr(autoprint = TRUE)


# for some reason having this plot first allows the brain plots to actually show up in the document. Not sure why, but going to keep this code in for now until I figure it out. 
xyz <- matrix(rnorm(27), ncol = 3)
triangles3d(xyz, col = rainbow(9))
spheres3d(xyz, col = rainbow(9), radius = 0.1)
```


```{r, test, webgl=TRUE, echo = FALSE, message=FALSE, warning=FALSE, results='hide'}
library(geomorph)

load(here::here('data/whole_brain_coords_S1.Rda'))
load(here::here('data/wholebrain_links.Rda'))

# GPA to get procrustes coordinates
wholebrain_gpa <- gpagen(wholebrain)

# PCA on procrustes coordinates
wholebrain_pca <- gm.prcomp(wholebrain_gpa$coords)



#rgl.snapshot('3dplot.png', fmt = 'png')


```


**Minimum PC1**
```{r, webgl = TRUE, echo = FALSE, message=FALSE, warning=FALSE}
# plotting the minimum and maximum specimen shape for PC1

plotRefToTarget(M1=wholebrain_pca$shapes$shapes.comp1$min, M2 = wholebrain_pca$shapes$shapes.comp1$min, method = 'points',mag = .5, links = wholebrain_links, axes = T, gridPar = gridPar(pt.size = 1))

```

**Maximum PC1**
```{r, max, webgl = TRUE, echo = FALSE, message=FALSE, warning=FALSE}

plotRefToTarget(M1=wholebrain_pca$shapes$shapes.comp1$max, M2 = wholebrain_pca$shapes$shapes.comp1$max, method = 'points',mag = .5, links = wholebrain_links, axes = T, gridPar = gridPar(pt.size = 1))

```



**Minimum PC2**
```{r, webgl = TRUE, echo = FALSE, message=FALSE, warning=FALSE}
# plotting the minimum and maximum specimen shape for PC1

plotRefToTarget(M1=wholebrain_pca$shapes$shapes.comp2$min, M2 = wholebrain_pca$shapes$shapes.comp2$min, method = 'points',mag = .5, links = wholebrain_links, axes = T, gridPar = gridPar(pt.size = 1))

```

**Maximum PC2**
```{r, webgl = TRUE, echo = FALSE, message=FALSE, warning=FALSE}

plotRefToTarget(M1=wholebrain_pca$shapes$shapes.comp2$max, M2 = wholebrain_pca$shapes$shapes.comp2$max, method = 'points',mag = .5, links = wholebrain_links, axes = T, gridPar = gridPar(pt.size = 1))

```

-----
Now trying different color palettes and comparing reference to target

Purple color palette - not as much a fan
**Maximum PC1**
```{r, webgl = TRUE, echo = FALSE, message=FALSE, warning=FALSE}

#M1 = reference
#M2 = target 
# pt.bg, link.col, pt.size = reference
# tar.pt.bg, tar.link.col, tar.pt.size = target

plotRefToTarget(M1= mshape(wholebrain_pca$shapes$shapes.comp1), 
                M2 = wholebrain_pca$shapes$shapes.comp1$max, 
                method = 'points',
                mag = .5, 
                links = wholebrain_links, 
                axes = T, 
                gridPar = gridPar(pt.bg = "thistle4",
                                  link.col = "thistle4",
                                  pt.size = 1, 
                                  tar.pt.bg = "mediumorchid4",
                                  tar.link.col = "mediumorchid4",
                                  tar.pt.size = 1))


```

Pink color palette - for now, I think this one is good  
Light pink = reference shape (average of all PC shapes)  
Dark pink = target shape (Maximum or Minimum PC)  
**Maximum PC1**
```{r, webgl = TRUE, echo = FALSE, message=FALSE, warning=FALSE}

#M1 = reference
#M2 = target 
# pt.bg, link.col, pt.size = reference
# tar.pt.bg, tar.link.col, tar.pt.size = target

plotRefToTarget(M1= mshape(wholebrain_pca$shapes$shapes.comp1), 
                M2 = wholebrain_pca$shapes$shapes.comp1$max, 
                method = 'points',
                mag = .5, 
                links = wholebrain_links, 
                axes = F, 
                gridPar = gridPar(pt.bg = "mistyrose3",
                                  link.col = "mistyrose3",
                                  pt.size = 1, 
                                  tar.pt.bg = "palevioletred4",
                                  tar.link.col = "palevioletred4",
                                  tar.pt.size = 0.75))


```

**Minimum PC1**
```{r, webgl = TRUE, echo = FALSE, message=FALSE, warning=FALSE}

#M1 = reference
#M2 = target 
# pt.bg, link.col, pt.size = reference
# tar.pt.bg, tar.link.col, tar.pt.size = target

plotRefToTarget(M1= mshape(wholebrain_pca$shapes$shapes.comp1), 
                M2 = wholebrain_pca$shapes$shapes.comp1$min, 
                method = 'points',
                mag = .5, 
                links = wholebrain_links, 
                axes = F, 
                gridPar = gridPar(pt.bg = "mistyrose3",
                                  link.col = "mistyrose3",
                                  pt.size = 1, 
                                  tar.pt.bg = "palevioletred4",
                                  tar.link.col = "palevioletred4",
                                  tar.pt.size = 0.75))


```


**Maximum PC2**
```{r, webgl = TRUE, echo = FALSE, message=FALSE, warning=FALSE}

#M1 = reference
#M2 = target 
# pt.bg, link.col, pt.size = reference
# tar.pt.bg, tar.link.col, tar.pt.size = target

plotRefToTarget(M1= mshape(wholebrain_pca$shapes$shapes.comp2), 
                M2 = wholebrain_pca$shapes$shapes.comp2$max, 
                method = 'points',
                mag = .5, 
                links = wholebrain_links, 
                axes = F, 
                gridPar = gridPar(pt.bg = "mistyrose3",
                                  link.col = "mistyrose3",
                                  pt.size = 1, 
                                  tar.pt.bg = "palevioletred4",
                                  tar.link.col = "palevioletred4",
                                  tar.pt.size = 0.75))


```

**Minimum PC2**
```{r, webgl = TRUE, echo = FALSE, message=FALSE, warning=FALSE}

#M1 = reference
#M2 = target 
# pt.bg, link.col, pt.size = reference
# tar.pt.bg, tar.link.col, tar.pt.size = target

plotRefToTarget(M1= mshape(wholebrain_pca$shapes$shapes.comp2), 
                M2 = wholebrain_pca$shapes$shapes.comp2$min, 
                method = 'points',
                mag = .5, 
                links = wholebrain_links, 
                axes = F, 
                gridPar = gridPar(pt.bg = "mistyrose3",
                                  link.col = "mistyrose3",
                                  pt.size = 1, 
                                  tar.pt.bg = "palevioletred4",
                                  tar.link.col = "palevioletred4",
                                  tar.pt.size = 0.75))


```










**Maximum PC3**
```{r, webgl = TRUE, echo = FALSE, message=FALSE, warning=FALSE}

#M1 = reference
#M2 = target 
# pt.bg, link.col, pt.size = reference
# tar.pt.bg, tar.link.col, tar.pt.size = target

plotRefToTarget(M1= mshape(wholebrain_pca$shapes$shapes.comp3), 
                M2 = wholebrain_pca$shapes$shapes.comp3$max, 
                method = 'points',
                mag = .5, 
                links = wholebrain_links, 
                axes = F, 
                gridPar = gridPar(pt.bg = "mistyrose3",
                                  link.col = "mistyrose3",
                                  pt.size = 1, 
                                  tar.pt.bg = "palevioletred4",
                                  tar.link.col = "palevioletred4",
                                  tar.pt.size = 0.75))


```

**Minimum PC3**
```{r, webgl = TRUE, echo = FALSE, message=FALSE, warning=FALSE}

#M1 = reference
#M2 = target 
# pt.bg, link.col, pt.size = reference
# tar.pt.bg, tar.link.col, tar.pt.size = target

plotRefToTarget(M1= mshape(wholebrain_pca$shapes$shapes.comp3), 
                M2 = wholebrain_pca$shapes$shapes.comp3$min, 
                method = 'points',
                mag = .5, 
                links = wholebrain_links, 
                axes = F, 
                gridPar = gridPar(pt.bg = "mistyrose3",
                                  link.col = "mistyrose3",
                                  pt.size = 1, 
                                  tar.pt.bg = "palevioletred4",
                                  tar.link.col = "palevioletred4",
                                  tar.pt.size = 0.75))


```




**Minimum PC3 vs Maximum PC3 and Min PC4 vs Max PC4**
```{r, webgl = TRUE, echo = FALSE, message=FALSE, warning=FALSE}

#M1 = reference
#M2 = target 
# pt.bg, link.col, pt.size = reference
# tar.pt.bg, tar.link.col, tar.pt.size = target

plotRefToTarget(M1= wholebrain_pca$shapes$shapes.comp3$min, 
                M2 = wholebrain_pca$shapes$shapes.comp3$max, 
                method = 'points',
                mag = .5, 
                links = wholebrain_links, 
                axes = F, 
                gridPar = gridPar(pt.bg = "mistyrose3",
                                  link.col = "mistyrose3",
                                  pt.size = 1, 
                                  tar.pt.bg = "palevioletred4",
                                  tar.link.col = "palevioletred4",
                                  tar.pt.size = 0.75))
# min = light pink
# max = dark pink

plotRefToTarget(M1= wholebrain_pca$shapes$shapes.comp4$min, 
                M2 = wholebrain_pca$shapes$shapes.comp4$max, 
                method = 'points',
                mag = .5, 
                links = wholebrain_links, 
                axes = F, 
                gridPar = gridPar(pt.bg = "mistyrose3",
                                  link.col = "mistyrose3",
                                  pt.size = 1, 
                                  tar.pt.bg = "palevioletred4",
                                  tar.link.col = "palevioletred4",
                                  tar.pt.size = 0.75))
```

**Maximum PC4**
```{r, webgl = TRUE, echo = FALSE, message=FALSE, warning=FALSE}

#M1 = reference
#M2 = target 
# pt.bg, link.col, pt.size = reference
# tar.pt.bg, tar.link.col, tar.pt.size = target

plotRefToTarget(M1= mshape(wholebrain_pca$shapes$shapes.comp4), 
                M2 = wholebrain_pca$shapes$shapes.comp4$max, 
                method = 'points',
                mag = .5, 
                links = wholebrain_links, 
                axes = F, 
                gridPar = gridPar(pt.bg = "mistyrose3",
                                  link.col = "mistyrose3",
                                  pt.size = 1, 
                                  tar.pt.bg = "palevioletred4",
                                  tar.link.col = "palevioletred4",
                                  tar.pt.size = 0.75))


```

**Minimum PC4**
```{r, webgl = TRUE, echo = FALSE, message=FALSE, warning=FALSE}

#M1 = reference
#M2 = target 
# pt.bg, link.col, pt.size = reference
# tar.pt.bg, tar.link.col, tar.pt.size = target

plotRefToTarget(M1= mshape(wholebrain_pca$shapes$shapes.comp4), 
                M2 = wholebrain_pca$shapes$shapes.comp4$min, 
                method = 'points',
                mag = .5, 
                links = wholebrain_links, 
                axes = F, 
                gridPar = gridPar(pt.bg = "mistyrose3",
                                  link.col = "mistyrose3",
                                  pt.size = 1, 
                                  tar.pt.bg = "palevioletred4",
                                  tar.link.col = "palevioletred4",
                                  tar.pt.size = 0.75))


```


export image rgl.snapshot('3dplot.png', fmt = 'png')
If error appear just change window from full screen to smaller screen and it will then work
